\pkg{NeuralNetTools}: Visualization and Analysis Tools for Neural Networks

### 2016-11-28 review_response.Rnw

I would recommend tentatively accepting the manuscript with the author addressing the following items:

$\bullet$ I would like to see a different applied example (or the current example changed) that gives a better practical demonstration of the ability to visualize a network and the corresponding interpretation.

{\it Content was added to the section that shows the interpretation of a neural network model using functions in the package.  Figure 7 was added to show the results from a simple model with an interpretation provided in the text.  Most of the addition to the text was included in the following paragraph (but see the applied example section for the entire revision):

``Figure 7 shows the information about arrival delays that can be obtained with the functions in \textbf{NeuralNetTools}. The NID (7a) shows the model structure and can be used as a general characterization of the relationships between variables.  For example, most of the connection weights from input node I5 are negative (grey), suggesting that distance travelled has an opposing relationship with arrival delays. Similarly, a large positive weight is observed between I3 and H1, suggesting arrival time is positively associated with arrival delays.  However, interpreting individual connection weights between layers is challenging.  Figures 7b amd 7c provide more quantitative descriptions using information from both the NID and model predictions.  Figure 7b shows variable importance using the \texttt{garson} and \texttt{olden} algorithms.  The \texttt{garson} function suggests time between destinations (\texttt{air\_time}) has the strongest relationship with arrival delays, similar to a strong positive association shown with the \texttt{olden} method.  However, the \texttt{garson} function shows arrival time (\texttt{arr\_time}) as the third most important variable for arrival delays, whereas this is shown as the strongest positive association by the \texttt{olden} function. Similar discrepancies between the two methods are observed for other variables, which are explained below. Finally, results from the \texttt{lekprofile} function (7c) confirm those in 7b, with the addition of non-linear responses that vary by different groupings of the data.  Values for each variable in the different unevaluated groups (based on clustering) show that there were no obvious patterns between groups with the exception being group four that generally had early arrival times and long departure delays.''    
}

@@

$\bullet$ Page 2, last paragraph:  NeuralNetTools does not ``improve the breadth and quality of information obtained from the MLP neural network.''  Instead, the package provides tools to better understand the MLP neural network.

{\it The sentence was revised: ''...that was developed to better understand information obtained from the MLP neural network.''}

$\bullet$ Page 6:  I am not familiar with the concepts of power and degrees of freedom with neural networks.  Please clarify the sentence, ``Pruning has additional benefits...''

{\it  These sentences were clarified to emphasize that pruning can aid in 1) the interpretation of relationships between variables by removing unimportant connections, and 2) successive model fitting by reducing the number of weights that are changed during model training: ``In addition to visualizing connections in the network that are not important, connections that are pruned can be removed in successive model fitting.  This reduces the number of free parameters (weights) that are  estimated by the model optimization routine, increasing the likelihood of convergence to an estimable numeric solution for the remaining connection weights that minimizes prediction error (i.e., model identifiability, Ellenius and Groth 2000).'' }

### 2016-11-28 cover_letter.tex
package license to cover letter
revision resubmitted to JSS

\begin{document}
\renewcommand{\rmdefault}{ptm}
\pagestyle{empty}

@@

\emph{To:}\newline
Bettina Gr\"un, Edzer Pebesma, Achim Zeileis\newline
Editors-in-Chief\newline
Journal of Statistical Software

@@

Enclosed please find my manuscript, entitled `NeuralNetTools: Visualization and Analysis Tools for Neural Networks', to be considered as a research article in the Journal of Statistical Software.

Supervised neural networks are statistical models that have been used to support prediction and exploratory analysis in data-rich environments.  A common criticism of these methods is that they represent `black-boxes' that do not describe any information about relationships among variables.  Methods to quantify information from a fitted neural network have been previously described in the literature but have not been available in a statistical computing environment.  This article describe the NeuralNetTools package for R that was developed to better quantify information provided by a neural network, including methods for plotting using a neural network interpretation diagram, estimates of variable importance, and a sensitivity analysis.  The package is exhaustive in scope by including methods for the most commonly used neural network objects in R.

Accompanying this submission is the manuscript PDF created with the JSS \LaTeX\ template, an R script of the manuscript code, and links to the package source code at the bottom of this letter.  Please note that the current version of the package on CRAN was submitted this year on Dec. 1 for minor revisions related to manuscript content.  The original package was first uploaded to CRAN on December 16\textsuperscript{th}, 2014.  Accordingly, the package has been used by the R community for a sufficient amount of time to address bugs or issues (see \href{https://github.com/fawda123/NeuralNetTools/issues?q=is%3Aissue+is%3Aclosed}{here}) in the early releases.

@@

\end{document}

>>>>

\begin{document}
\renewcommand{\rmdefault}{ptm}
\pagestyle{empty} 
\small

@@

\vfill
\emph{To:}\newline
Bettina Gr\"un, Edzer Pebesma, Achim Zeileis\newline
Editors-in-Chief\newline
Journal of Statistical Software

@@

Enclosed please find my revised manuscript, entitled `NeuralNetTools: Visualization and Analysis Tools for Neural Networks'.  This revision follows a minor acceptence as a research article in JSS.

I appreciate the insightful comments provided by the reviewer and I have made every effort to address the concerns.  In particular, I have improved the applied example by demonstrating the interpretation of a simple neural network model using functions from the package. This includes addition to the text and Figure 7.  I have also fixed any remaining issues with reproducibility in the manuscript code and supplementary material.  Please see all of my responses to the reviewer comments included with my revision.

This revision includes a manuscript PDF created with the JSS \LaTeX\ template (nnt\_manu.pdf), my response to the reviewer comments (review\_response.pdf), an R script of the manuscript code (nnt\_manu.R), a .tex file for the manuscript text (nnt\_manu.tex), a .bbl file with bibliographic information (refs.bbl), all pdf files used for figures (in fig folder), and supporting JSS \LaTeX\ template files.  Links to the package source code are at the bottom of this letter. The package is licensed uncer CC0.

@@

\vfill
\end{document}

### 2016-11-28 nnt_manu.Rnw
revision resubmitted to JSS

The \pkg{RSNNS} package provides algorithms to prune connections or nodes in a neural network \citep{Bergmeir12}.  This approach can remove connection weights between layers or input nodes that do not contribute to the predictive performance of the network.  In addition to visualizing connections in the network that are not important, connections that are pruned can be removed in successive model fitting.  This reduces the number of free parameters (weights) that are  estimated by the model optimization algorithm, thereby increasing the likelihood of convergence to an estimable numeric solution for the remaining connection weights that minimizes prediction error \citep[i.e., model identifiability,][]{Ellenius00}. Algorithms in \pkg{RSNNS} for weight pruning include magnitude-based pruning, Optimal Brain Damage, and Optimal Brain Surgeon, whereas algorithms for node pruning include skeletonization and the non-contributing units method \citep{Zell98}.  The \code{plotnet} function can plot a pruned neural network, with options to omit or display the pruned connections (\cref{fig:plotprune}).

@@

\Cref{fig:flightplots} shows the information about arrival delays that can be obtained with the functions in \pkg{NeuralNetTools}. The \ac{nid} (\ref{fig:flightplotsa}) shows the model structure and can be used develop a general characterization of the relationships between variables.  For example, most of the connection weights from input node I5 are negative (grey), suggesting that distance travelled has an opposing relationship with arrival delays. \Cref{fig:flightplotsb} and \ref{fig:flightplotsc} provide more quantitative descriptions using information from both the \ac{nid} and model predictions.  \Cref{fig:flightplotsb} shows variable importance using the \code{garson} and \code{olden} algorithms.  The \code{garson} function suggests time between destinations (\code{air\_time}) has the strongest relationship with arrival delays, similar to a strong positive association shown with the \code{olden} method.  However, the \code{garson} function shows arrival time as the second most important variable, whereas the \code{olden} function shows this variable as having a weak negative relationship with arrival time. This discrepancy is explained below. Finally, results from the \code{lekprofile} function (\ref{fig:flightplotsc}) confirm those in \ref{fig:flightplotsb}, with the addition of non-linear responses that vary by different groupings of the data.  Values for each variable in the different unevaluated groups (based on clustering) show that there were no obvious patterns between groups with the exception being group four that generally had early arrival times and long departure delays.

>>>>

The \pkg{RSNNS} package provides algorithms to prune connections or nodes in a neural network \citep{Bergmeir12}.  This approach can remove connection weights between layers or input nodes that do not contribute to the predictive performance of the network.  In addition to visualizing connections in the network that are not important, connections that are pruned can be removed in successive model fitting.  This reduces the number of free parameters (weights) that are  estimated by the model optimization algorithm, increasing the likelihood of convergence to an estimable numeric solution for the remaining connection weights that minimizes prediction error \citep[i.e., model identifiability,][]{Ellenius00}. Algorithms in \pkg{RSNNS} for weight pruning include magnitude-based pruning, Optimal Brain Damage, and Optimal Brain Surgeon, whereas algorithms for node pruning include skeletonization and the non-contributing units method \citep{Zell98}.  The \code{plotnet} function can plot a pruned neural network, with options to omit or display the pruned connections (\cref{fig:plotprune}).

@@

\Cref{fig:flightplots} shows the information about arrival delays that can be obtained with the functions in \pkg{NeuralNetTools}. The \ac{nid} (\ref{fig:flightplotsa}) shows the model structure and can be used as a general characterization of the relationships between variables.  For example, most of the connection weights from input node I5 are negative (grey), suggesting that distance travelled has an opposing relationship with arrival delays. Similarly, a large positive weight is observed between I3 and H1, suggesting arrival time is positively associated with arrival delays.  However, interpreting individual connection weights between layers is challenging.  \Cref{fig:flightplotsb} and \ref{fig:flightplotsc} provide more quantitative descriptions using information from both the \ac{nid} and model predictions.  \Cref{fig:flightplotsb} shows variable importance using the \code{garson} and \code{olden} algorithms.  The \code{garson} function suggests time between destinations (\code{air\_time}) has the strongest relationship with arrival delays, similar to a strong positive association shown with the \code{olden} method. However, the \code{garson} function shows arrival time (\code{arr\_time}) as the third most important variable for arrival delays, whereas this is shown as the strongest positive association by the \code{olden} function. Similar discrepancies between the two methods are observed for other variables, which are explained below. Finally, results from the \code{lekprofile} function (\ref{fig:flightplotsc}) confirm those in \ref{fig:flightplotsb}, with the addition of non-linear responses that vary by different groupings of the data.  Values for each variable in the different unevaluated groups (based on clustering) show that there were no obvious patterns between groups with the exception being group four that generally had early arrival times and long departure delays.

### 2016-11-28 resubmit/nnt_manu.tex
revision resubmitted to JSS

\Cref{fig:flightplots} shows the information about arrival delays that can be obtained with the functions in \pkg{NeuralNetTools}. The \ac{nid} (\ref{fig:flightplotsa}) shows the model structure and can be used develop a general characterization of the relationships between variables.  For example, most of the connection weights from input node I5 are negative (grey), suggesting that distance travelled has an opposing relationship with arrival delays. \Cref{fig:flightplotsb} and \ref{fig:flightplotsc} provide more quantitative descriptions using information from both the \ac{nid} and model predictions.  \Cref{fig:flightplotsb} shows variable importance using the \code{garson} and \code{olden} algorithms.  The \code{garson} function suggests time between destinations (\code{air\_time}) has the strongest relationship with arrival delays, similar to a strong positive association shown with the \code{olden} method.  However, the \code{garson} function shows arrival time as the second most important variable, whereas the \code{olden} function shows this variable as having a weak negative relationship with arrival time. This discrepancy is explained below. Finally, results from the \code{lekprofile} function (\ref{fig:flightplotsc}) confirm those in \ref{fig:flightplotsb}, with the addition of non-linear responses that vary by different groupings of the data.  Values for each variable in the different unevaluated groups (based on clustering) show that there were no obvious patterns between groups with the exception being group four that generally had early arrival times and long departure delays.

@@

% note that ref titles need to be in title-case 
% \bibliographystyle{jss}
% \bibliography{refs}
% this file is bbl edited by hand after running bibtex
\input{CopyOfnnt_manu.bbl}

>>>>

\Cref{fig:flightplots} shows the information about arrival delays that can be obtained with the functions in \pkg{NeuralNetTools}. The \ac{nid} (\ref{fig:flightplotsa}) shows the model structure and can be used as a general characterization of the relationships between variables.  For example, most of the connection weights from input node I5 are negative (grey), suggesting that distance travelled has an opposing relationship with arrival delays. Similarly, a large positive weight is observed between I3 and H1, suggesting arrival time is positively associated with arrival delays.  However, interpreting individual connection weights between layers is challenging.  \Cref{fig:flightplotsb} and \ref{fig:flightplotsc} provide more quantitative descriptions using information from both the \ac{nid} and model predictions.  \Cref{fig:flightplotsb} shows variable importance using the \code{garson} and \code{olden} algorithms.  The \code{garson} function suggests time between destinations (\code{air\_time}) has the strongest relationship with arrival delays, similar to a strong positive association shown with the \code{olden} method. However, the \code{garson} function shows arrival time (\code{arr\_time}) as the third most important variable for arrival delays, whereas this is ranked highest by the \code{olden} function. Similar discrepancies between the two methods are observed for other variables, which are explained below. Finally, results from the \code{lekprofile} function (\ref{fig:flightplotsc}) confirm those in \ref{fig:flightplotsb}, with the addition of non-linear responses that vary by different groupings of the data.  Values for each variable in the different unevaluated groups (based on clustering) show that there were no obvious patterns between groups with the exception being group four that generally had early arrival times and long departure delays.

@@

% note that ref titles need to be in title-case 
\input{refs.bbl}
